---
title: "Attrition RMD"
author: "Samuel"
date: "8/4/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Summary
#The purpose of this project is to explore what variables are good predictors for attrition rates in Fortune 1000 companies and also we will create a model that will predict if an employee will leave their company voluntarily or not. Finally, we will look at other trends associated with specific jobs and attrition rates.
```

```{r}
#environment libraries
library(tidyverse)
library(corrplot)
library(mlbench)
library(caret)
library(skimr)
library(mice)
library(purrr)
library(ggplot2)
library(ggthemes)
library(cowplot)
library(class)
library(e1071)
library(grid)
library(tidyr)
library(stringr)
library(naniar)

```

# Build a model for Attrition 

# Start by Tidying and Exploring the dT
#I am cheking for the Correlation, the Trends and the Causation

# Remove missing values if available in data set
# Seperate continous variables from categorical variables
# Remove  unwnted variables remove from the list
# Check and make sure categorical variables are stored as factors.
# Remove  highly correlated numeric variables as they tends to interpret the same message
# Perform EDA on  Numeric vs Categorical
# Perform EDA on  Categorical vs Categorical
# KNN Model V

```{r}
dfTrain <- read.csv("CaseStudy2-data.csv", header = TRUE, strip.white=TRUE)
dfTrain
head(dfTrain)
skim(dfTrain)
length(dfTrain)
invisible(ncol(dfTrain))
invisible(skim(dfTrain))
```



# Job Role Trends
## Categorical v Categorical
- After reviewing the categorical variables to compare to our Job Roles the below would be insightful for our clients
  - Attrition, OverTime, Job Satisfaction

```{r}
#str(dfTrain)

JobdfTrain = dfTrain
skim(JobdfTrain)

JobdfTrain$Education <- as.factor(dfTrain$Education)
JobdfTrain$EnvironmentSatisfaction <- as.factor(dfTrain$EnvironmentSatisfaction)
JobdfTrain$JobLevel <- as.factor(dfTrain$JobLevel)
JobdfTrain$JobInvolvement <- as.factor(dfTrain$JobInvolvement)
JobdfTrain$JobSatisfaction <- as.factor(dfTrain$JobSatisfaction)
JobdfTrain$RelationshipSatisfaction <- as.factor(dfTrain$RelationshipSatisfaction)
JobdfTrain$StockOptionLevel <- as.factor(dfTrain$StockOptionLevel)
JobdfTrain$WorkLifeBalance <- as.factor(dfTrain$WorkLifeBalance)

skim(JobdfTrain)
ncol(JobdfTrain)
#JobdfTrain %>% ggplot() + geom_bar(aes(x = Attrition, fill = JobRole), position = "fill", alpha = 0.9) + coord_flip()

#JobdfTrain %>% ggplot() + geom_bar(aes(x = Gender, fill = JobRole), position = "fill", alpha = 0.9) + coord_flip()
#JobdfTrain %>% ggplot() + geom_bar(aes(x = JobSatisfaction, fill = JobRole), position = "fill", alpha = 0.9) + coord_flip()
#JobdfTrain %>% ggplot() + geom_bar(aes(x = OverTime, fill = JobRole), position = "fill", alpha = 0.9) + coord_flip()
#JobdfTrain %>% ggplot() + geom_bar(aes(x = WorkLifeBalance, fill = JobRole), position = "fill", alpha = 0.9) + coord_flip()

#Remove variables that show less trends

#Funtion to plot a grip barplot for  job role trends
# Name target variable
JobroleVar <- "JobRole"

# Name explanatory variable
ExpVar <- JobdfTrain %>% keep(is.factor) %>% colnames

# Create function
JobroleFunc <- function(df, explanatory, response) {
  ggplot(data = df) + geom_bar(aes_string(x = explanatory, fill = response), position = "fill", alpha = 0.9) + coord_flip() + xlab(explanatory) + ylab("")
}

  # Example of working function above
  #JobroleFunc(dfTrain, explanatory = "Education", response = "Attrition")


#  Create plot list for plot_grid function to reference
JobrolePlotList <- lapply(ExpVar, function(x) JobroleFunc(JobdfTrain, x, JobroleVar))

#  Grid of all categorical variables plotted against Attrition
plot_grid(plotlist = JobrolePlotList)
```

## Job role Categorical v Numeric
- After reviewing the numeric variables to compare to our Job Roles the below would be insightful for our clients
  - Age, Years Since Last Promotion, Distance from Home, Years At Company

```{r Job Role v Age}
dfTrain %>% ggplot() + geom_boxplot(aes(x = JobRole, y = Age, fill = JobRole)) +
  theme_excel_new() + ggtitle("Job Role v Age") + ylab("Education")
```

```{r Job Role v Years Since Last Promotion}
dfTrain %>% ggplot() + geom_boxplot(aes(x = JobRole, y = YearsSinceLastPromotion, fill = JobRole)) +
  theme_excel_new() + ggtitle("Job Role v Years Since Last Promotion")
```

```{r Job Role v Distance from Home}
dfTrain %>% ggplot() + geom_boxplot(aes(x = JobRole, y = DistanceFromHome, fill = JobRole)) +
  theme_excel_new() + ggtitle("Job Role v Distance from Home")
```

```{r Job Role v Years At Company}
dfTrain %>% ggplot() + geom_boxplot(aes(x = JobRole, y = YearsAtCompany, fill = JobRole)) +
  theme_excel_new() + ggtitle("Job Role v Years At Company")
```


```{r}
#Continuous variables are: 
# "TotalWorkingYears", "TrainingTimesLastYear", "YearsAtCompany","Age", , "HourlyRate", "MonthlyIncome", "MonthlyRate", "NumCompaniesWorked", , "YearsSinceLastPromotion", "YearsWithCurrManager", "StandardHours" "PercentSalaryHike",  "YearsInCurrentRole", "DailyRate", "DistanceFromHome", "EmployeeCount"

#Categorical Variables are:
# "Attrition","Over18" , "MaritalStatus",  "JobRole", "OverTime""EducationField", "Gender", "Department", "BusinessTravel",

# Variabless with Level
# "StockOptionLevel", "WorkLifeBalance" "Education", "JobLevel",  "JobInvolvement", "EnvironmentSatisfaction", "JobSatisfaction", "PerformanceRating", "RelationshipSatisfaction"


```


# Missing data evaluation {data-background=""}
There is no missing value

```{r NA eval}
vis_miss(dfTrain)
md.pattern(dfTrain)

#### We see that there are 3 patterns: 
 # 870  observations with complete information 
#  0 observations have a NA in IBU
# 32 variables

```

# Remove all zero variance variables, this can be checked using the skim() function

```{r zero variance variables}
invisible(skim(dfTrain))
dfTrain2 <- select(dfTrain, -c("EmployeeCount", "Over18", "StandardHours"))
skim(dfTrain2)
ncol(dfTrain2)
```


# Remove unwanted variables

```{r}

dfTrain3 <- select(dfTrain2, -c("EmployeeNumber", "ID", "PerformanceRating"))
skim(dfTrain3)
ncol(dfTrain3)

```

# Check if all categorical variables are stored as factors, if not then convert all categorical variables to factors
# The categorical variables are; "Attrition", "BusinessTravel", "Department", "EducationField", "Gender", "JobRole", "MaritalStatus", "Over18", "OverTime"


# Storing all level numeric variables as factors

```{r level variables}
# Categorical Variables
skim(dfTrain3)

dfTrain3$Education <- as.factor(dfTrain3$Education)
dfTrain3$EnvironmentSatisfaction <- as.factor(dfTrain3$EnvironmentSatisfaction)
dfTrain3$JobLevel <- as.factor(dfTrain3$JobLevel)
dfTrain3$JobInvolvement <- as.factor(dfTrain3$JobInvolvement)
dfTrain3$JobSatisfaction <- as.factor(dfTrain3$JobSatisfaction)
dfTrain3$RelationshipSatisfaction <- as.factor(dfTrain3$RelationshipSatisfaction)
dfTrain3$StockOptionLevel <- as.factor(dfTrain3$StockOptionLevel)
dfTrain3$WorkLifeBalance <- as.factor(dfTrain3$WorkLifeBalance)

skim(dfTrain3)
ncol(dfTrain3)
```

# Checking to see the correlation on numeric variables
#Highly correlated relationships
#Total Working Years v Monthly Income: 78%
#Years at Company v Years in Current Role: 78%
#Years at Company v Years with Current Manager: 77%
#Years at Company v Total Working Years: 64%
#Years at Company v Years Since Last Promotion: 64%
#Years with Current Manager v Years in Current Role: 71%
#Age vs Tota Working Years: 65%
  

```{r high correlation}
corrdfTrain <- dfTrain3
# This table shows the correlation between the numerical variables
corrdfTraintable <- corrdfTrain %>% keep(is.numeric) %>% na.omit %>% cor %>% view

corrdfTrain %>% keep(is.numeric) %>% na.omit %>% cor %>% corrplot("upper", addCoef.col = "black", number.digits = 2, number.cex = 0.5, method="shade", order="hclust", tl.srt=45, tl.cex = 0.8)
invisible(view(corrdfTraintable))
```

# Correlation Review
- Reviewing the correlated variables we can logically conclude some of these are saying similar things. However, the high correlation doesn't tell the whole story. We will keep investigating to get a better picture of these variables and their importance to Attrition.
  - Keep: "MonthlyIncome", "YearsInCurrentRole", "YearsSinceLastPromotion", "YearsAtCompany", "TotalWorkingYears", "YearsWithCurrManager"
  - Drop: "YearsatCompany"
  
```{r}
skim(dfTrain3)

dfTrain3 <- select(dfTrain3, -c("YearsAtCompany"))
skim(dfTrain3)
ncol(dfTrain3)
```

# EDA Numerical variables vs Categorical variables
#Comparison of Attrition (the variable of interest) to categorical variables left in model
# Plot will be filled by Attrition
#If density plot shows peaks in separate areas then important variable
#If density plot shows peaks in the same spot (clean overlay) then not an important variable

```{r Automated EDA: Numeric v  Catgorical: all at once}
# 1. Name target variable
Attr <- "Attrition"

# 2. Name explanatory variable
ExplVar<- dfTrain3 %>% keep(is.numeric) %>% colnames

# 3. Create function
PlotFunc <- function(df, explanatory, response) {
  ggplot(data = df) + geom_density(aes_string(x = explanatory, fill = response), alpha = 0.5) + xlab(explanatory) + ylab("Attrition")
}

  # Density plot
PlotFunc(dfTrain3, explanatory =  "TotalWorkingYears", response = "Attrition")

# 4. Create plot list for plot_grid function to reference
PlotList <- lapply(numvariables, function(x) PlotFunc(dfTrain3, x, Attr))

#  Grid of all categorical variables plotted against Attrition
plot_grid(plotlist = PlotList)
```

# After Numeric v Categorical Review
# Get rid of variables that have similar trends
# Remove variables that show the Yes and No of Attrition having the same patterns
#Keep variables that are showing opposite peaks as in the graphs
#"MonthlyIncome", "YearsInCurrentRole", "NumCompaniesWorked", "Age"

```{r}

dfTrain4 <- select(dfTrain3, -c("TrainingTimesLastYear", "PercentSalaryHike", "DistanceFromHome", "DailyRate", "HourlyRate", "TotalWorkingYears", "MonthlyRate", "YearsWithCurrManager", "YearsSinceLastPromotion"))
skim(dfTrain4)
ncol(dfTrain4)
```


### 8. Categorical v Categorical (Attrition)
#Comparison of Attrition (dependent variable of interest) to categorical variables in model
#Bar chart with percentages instead of totals
#Colored by Attrition
#If bar chart shows a large difference within the levels/categories of the y variable then it is identified as an important variable to include in the model
#If bar chart shows a small or zero difference within the levels/categories of the y variable then it is not an important variable to include in the model

```{r}
str(dfTrain4)
dfTrain4 %>% ggplot() + geom_bar(aes(x = Education, fill = Attrition), position = "fill", alpha = 0.9) + coord_flip()
```


#### Automated EDA: Categorical v Categorical (Attrition) all at once

```{r}
# 1. Name target variable
FuncAttr <- "Attrition"

# 2. Name explanatory variable
Func2 <- dfTrain4 %>% keep(is.factor) %>% colnames

# 3. Create function
BarFunc <- function(df, explanatory, response) {
  ggplot(data = df) +geom_bar(aes_string(x = explanatory, fill = response), position = "fill", alpha = 0.9) + coord_flip() + xlab(explanatory)
}

  # Example of working function above
  #BarFunc(dfTrain4, explanatory = "MaritalStatus", response = "Attrition")


# 4. Create plot list for plot_grid function to reference
plotlist2 <- lapply(Func2, function(x) BarFunc(dfTrain3, x, FuncAttr))

# 5. Grid of all categorical variables plotted against Attrition
plot_grid(plotlist = plotlist2)
```

# Categorical v Categorical Review
# Drop variables that show a small difference within the levels/categories of the response variable and Keep variables that show a large difference within the levels/categories

```{r}
#skim(dfTrain4)

dfTrain5 <- select(dfTrain4, -c("BusinessTravel", "Department", "Education", "EducationField", "EnvironmentSatisfaction", "Gender", "JobRole", "JobSatisfaction", "MaritalStatus", "RelationshipSatisfaction", "JobLevel"))
#skim(dfTrain5)
ncol(dfTrain5)

#view(dfTrain5)
#str(dfTrain5)
dfTrain6 <- dfTrain5 %>% 
  mutate(OverTime = ifelse(OverTime == "No", 0, 1))
dfTrain6 <- dfTrain5 %>%
  mutate(Attrition = ifelse(Attrition ==  "No", 0, 1))
#view(dfTrain6)
```



#Let's narrow down on the variables by using KNN 
- Categorical: "JobInvolvement", "JobLevel", "OverTime", "StockOptionLevel", "WorkLifeBalance"
- Numerical: "Age", "MonthlyIncome", "NumCompaniesWorked", "YearsInCurrentRole"

### Review of multiple models

# Age, Job Involvement, Job Level

```{r}
#invisible(knn(traindfTrain[,c(1,3,4)], testdfTrain[,c(1,3,4)], traindfTrain$Attrition, prob = TRUE, k=19))
#classificationdfTrain = knn(traindfTrain[,c(1,3,4)], testdfTrain[,c(1,3,4)], traindfTrain$Attrition, prob = TRUE, k=19)
#CMdfTrain1.1 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
#CMdfTrain1.1
```




# Job Involvement, Monthly Income, Number Companies Worked

```{r Job Involvement, Monthly Income, Number Companies Worked}
#invisible(knn(traindfTrain[,c(3,4,5)], testdfTrain[,c(3,4,5)], traindfTrain$Attrition, prob = TRUE, k=19))
#classificationdfTrain = knn(traindfTrain[,c(3,4,5)], testdfTrain[,c(3,4,5)], traindfTrain$Attrition, prob = TRUE, k=19)
#CMdfTrain2 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
#CMdfTrain2
```

# Monthly Income, Number of Companies Worked, Over Time

```{r Monthly Income, Number of Companies Worked}
#invisible(knn(traindfTrain[,c(4,5,6)], testdfTrain[,c(4,5,6)], traindfTrain$Attrition, prob = TRUE, k=7))
#classificationdfTrain = knn(traindfTrain[,c(4,5,6)], testdfTrain[,c(4,5,6)], traindfTrain$Attrition, prob = TRUE, k=7)
#CMdfTrain3 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
#CMdfTrain3
```

# Number of Companies Worked, Over Time, Stock Option Level

```{r Number of Companies Worked, Over Time, Stock Option Level}
#invisible(knn(traindfTrain[,c(5,6,7)], testdfTrain[,c(5,6,7)], traindfTrain$Attrition, prob = TRUE, k=7))
#classificationdfTrain = knn(traindfTrain[,c(5,6,7)], testdfTrain[,c(5,6,7)], traindfTrain$Attrition, prob = TRUE, k=7)
#CMdfTrain5 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
#CMdfTrain5
```

# Over Time, WorkLifeBalance, Years In Current Role

```{r}
#invisible(knn(traindfTrain[,c(6,7,8)], testdfTrain[,c(6,7,8)], traindfTrain$Attrition, prob = TRUE, k=7))
#classificationdfTrain = knn(traindfTrain[,c(6,7,8)], testdfTrain[,c(6,7,8)], traindfTrain$Attrition, prob = TRUE, k=7)
#CMdfTrain6 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
#CMdfTrain6
```


# Stock Option Level, Work Life Balance, Age

```{r}
#invisible(knn(traindfTrain[,c(7,8,1)], testdfTrain[,c(7,8,1)], traindfTrain$Attrition, prob = TRUE, k=19))
#classificationdfTrain = knn(traindfTrain[,c(7,8,1)], testdfTrain[,c(7,8,1)], traindfTrain$Attrition, prob = TRUE, k=19)
#CMdfTrain7 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
#CMdfTrain7
```

# Work Life Balance, Years in Current Role, Age

```{r}
#invisible(knn(traindfTrain[,c(8,1,3)], testdfTrain[,c(8,1,3)], traindfTrain$Attrition, prob = TRUE, k= 19))
#classificationdfTrain = knn(traindfTrain[,c(8,1,3)], testdfTrain[,c(8,1,3)], traindfTrain$Attrition, prob = TRUE, k= 19)
#CMdfTrain8 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
#CMdfTrain8
```

# Montly Income (4) Over Time(6) Stock Option Level (7)
```{r}
#invisible(knn(traindfTrain[,c(4,6,7)], testdfTrain[,c(4,6,7)], traindfTrain$Attrition, prob = TRUE, k=19))
#classificationdfTrain = knn(traindfTrain[,c(4,6,7)], testdfTrain[,c(4,6,7)], traindfTrain$Attrition, prob = TRUE, k=19)
#CMdfTrain8 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
#CMdfTrain8
```



#### Find best k for Model 7
- Best k: 15 - 23
```{r find best k}
set.seed(8)
splitPerc = .8
trnInddfTrain = sample(1:dim(dfTrain6)[1], round(splitPerc * dim(dfTrain6)[1]))

traindfTrain = dfTrain6[trnInddfTrain,]

testdfTrain = dfTrain6[-trnInddfTrain,]

spec = data.frame(specificity = numeric(30), k = numeric(30))
for (i in 1:30) 
  {
  classificationdfTrain = knn(traindfTrain[,c(8,1,3)], testdfTrain[,c(8,1,3)], traindfTrain$Attrition, prob = TRUE, k=19)
  table(testdfTrain$Attrition, classificationdfTrain)
  CMdfTrain8 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
  spec$specificity[i] = CMdfTrain8$byClass[2]
  spec$k[i] = i
}
plot(spec$k, spec$specificity, type = "l", xlab = "k")
#k <- which.max(spec$specificity)
```


#### Final Model (7)

```{r}
invisible(knn(traindfTrain[,c(8,1,3)], testdfTrain[,c(8,1,3)], traindfTrain$Attrition, prob = TRUE, k= 20))
classificationdfTrain = knn(traindfTrain[,c(8,1,3)], testdfTrain[,c(8,1,3)], traindfTrain$Attrition, prob = TRUE, k= 20)
CMdfTrain8 = confusionMatrix(table(testdfTrain$Attrition, classificationdfTrain))
CMdfTrain8
```

## Final KNN Prediction Model for Attrition
# Attrition is most closely identified by taking into account
# Number of companies a person has worked.
# If they do or do not have to work overtime.
# An employee's stock option level
# Accuracy > 84%
# Sensitivity > 84%
# Specificity = 75%


## CompSet Set Analysis
```{r}
CompSet <- read.csv("CaseStudy2CompSet No Attrition.csv", header = TRUE, strip.white=TRUE)
head(CompSet)
```

## KNN Model Used to Predict Attrition
```{r}
#Work Life Balance, Years in Current Role, Age
CompSet <- select(CompSet, c("ID", "WorkLifeBalance", "YearsInCurrentRole", "Age"))

estimates <- data.frame(knn(traindfTrain[,c(8,1,3)], CompSet[,c(2,3,4)], traindfTrain$Attrition, prob = TRUE, k = 19))

#Join the original table and the competition set
CompSetResult <- cbind(CompSet, estimates)

#Set Column name for Attrition
colnames(CompSetResult)[5] <- "Attrition"

CompSetResult <- CompSetResult %>% 
  mutate(Attrition = ifelse(Attrition == 0, "No", "Yes"))
invisible(CompSetResult)
```

### Checking model results aginst graphical analysis

```{r}
CompSetResult %>% ggplot(mapping = aes(x = Attrition)) + geom_bar() + ggtitle("Comeptition Set: Attrition Count") + theme_excel_new()

dfTrain %>% ggplot(mapping = aes(x = Attrition)) + geom_bar() + ggtitle("Original Data: Attrition Count") + theme_excel_new()
```

### Checking model results aginst graphical analysis

```{r}
CompSetResult %>% ggplot() + geom_density(aes(x = Age, fill = Attrition, alpha = .1)) +
  theme_excel_new() + ggtitle("Attrition v Age") + ylab("Attrition") + xlab("Age")
```

### Checking model against graphical analysis

```{r}
CompSetResult %>% ggplot() + geom_bar(aes(x = WorkLifeBalance, fill = Attrition), position = "fill", alpha = .7) + coord_flip()
```

### Write CSV of fitted CompSet Data Set
```{r}
#write_csv(CompSetResult, "/Users/olusho/Documents/CaseStudy2DDS\\CCase2Predictions  AttritionOnalaja.csv")
```


# Salary set analysis
## KNN Model Used to Predict Attrition
```{r}

SalarySet <- read.csv("CaseStudy2CompSet No Salary.csv", header = TRUE, strip.white=TRUE)
head(SalarySet)

#Work Life Balance, Years in Current Role, Age
Salary <- select(SalarySet, c("ID", "WorkLifeBalance", "YearsInCurrentRole", "Age"))

SalaryEst <- data.frame(knn(traindfTrain[,c(8,1,3)], Salary[,c(2,3,4)], traindfTrain$MonthlyIncome, prob = TRUE, k = 19))

#Join the original table and the competition set
SalSetResult <- cbind(Salary, SalaryEst)

#Set Column name for Attrition
colnames(SalSetResult)[5] <- "Salary"

write_csv(SalSetResult, "/Users/olusho/Documents/CaseStudy2DDS\\CCase2Predictions  SalaryOnalaja.csv")

#RMSE

set.seed(3)
splitPerc = .8

DimSalary = sample(1:dim(SalSetResult)[1], round(splitPerc * dim(SalSetResult)[1]))
trainSalary = SalSetResult[DimSalary,]
testSalary = SalSetResult[-DimSalary,]

a = as.numeric(SalSetResult$Salary)
SalaryModel = lm(a ~ WorkLifeBalance + YearsInCurrentRole + Age, data = SalSetResult)
summary(SalaryModel)

PredSalaryModel = predict(SalaryModel, newdata = testSalary)

b = as.numeric(testSalary$Salary)

RSME1 = sqrt(mean((b - PredSalaryModel)^2))










```


